---
title: "Practical-Machine-Learning-JHU-Final-Project"
author: "Faustin GASHAKAMBA"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Assignment Overview

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, I will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants who were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The goal will be to predict the manner in which these 6 participants did the exercise. To achieve this, I will use a number of predictors among the 20 variables available in the data to predict the outcome "classe". I will then create a report describing how I built the model, how I performed cross validation, what the expected out of sample error is, and why I made the choices I did.

STEP 1: Installing and loading the necessary packages
```{r caret, echo = TRUE}
library(caret)
```

```{r rpart, echo = TRUE}
library(rpart)
library(rpart.plot)
```

```{r ColorBrewer, echo=TRUE}
library(RColorBrewer)
library(rattle)
```

```{r rf, echo = TRUE}
library(randomForest)
```

STEP 2: Setting the seed for results reproducibility
```{r seed, echo = TRUE}
set.seed(12345)
```

STEP 3: Loading data
```{r Data, echo = TRUE}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
dim(training)
dim(testing)
```

STEP 4: Cleaning the training dataset
1: Cleaning NearZeroVariance Variables
```{r NZV, echo = TRUE}
myDataNZV <- nearZeroVar(training, saveMetrics=TRUE)
myNZVvars <- names(training) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
training <- training[!myNZVvars]
dim(training)
```

2: Remove the ID column
```{r ID, echo = TRUE}
training <- training[c(-1)]
```

3: Updating NA values
```{r NA, echo = TRUE}
trainingV3 <- training 
for(i in 1:length(training)) { 
if( sum( is.na( training[, i] ) ) /nrow(training) >= .6 ) { 
for(j in 1:length(trainingV3)) {
if( length( grep(names(training[i]), names(trainingV3)[j]) ) ==1)  { 
trainingV3 <- trainingV3[ , -j] #Remove that column
}
}
}
}
training <- trainingV3
rm(trainingV3)
```

STEP 5: Partioning the training set into two
```{r partitionning, echo = TRUE}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
```

STEP 6: Train a model with Decision Trees
```{r lta, echo = TRUE}
set.seed(12345)
rf_Fit <- train(classe ~ ., data=myTraining, method="rpart")
#fancyRpartPlot(rf_Fit)
```

STEP 7: Prediction with Decision Trees Model
```{r predict, echo = TRUE}
pred_Rf <- predict(rf_Fit, newdata = myTesting)
conf_tree <- confusionMatrix(pred_Rf, as.factor(myTesting$classe))
conf_tree
```